---
title: "Database-detections"
output: rmarkdown::pdf_document
vignette: >
  %\VignetteIndexEntry{Database-detections}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = FALSE,
  echo = TRUE,
  comment = "#>"
)
```

This is a subtopic of the `Database` workflow.  It covers all handling of the detection data, from the files downloaded in the field to appending the database.

## General detections workflow

  1. Download acoustic receivers in the field
  2. Copy all vrls from a given field download range into a new subfolder of `YB_Vemco/` using the `ybt::copy_vrls()` function.
  3. Create a new .vdb in that subfolder, following the naming conventions below.
  4. Process data in VUE, including drift-correction and exporting to .csv
  5. Format the detection .csv(s) in R with the `ybt::format_VUE_detections()` function
  6. Append the resulting detections dataframe to SQLite database using the `ybt::ybt_db_append()` function
  7. Back up updated database to cloud and/or external storage

## 1. Download Receivers in the field

 - Create a VUE database (.vdb) for field downloads
 
     - Filename convention: "YB_fieldyyyy_mmdd.vdb"
        - for example: "YB_field2020_0321.vdb" for a download occurring on March 21st 2020.
     
    - Put the .vdb, along with each individual .vrl and its RLD file, into its own subdirectory of the "FieldDownloads" folder.  Example:

```{r, echo=TRUE, eval=FALSE}
YB_Vemco
 |-- YBFieldDownloads
 |    |-- Field2020_0901
 |        `--Field2020_0901.vdb
 |        `--VR2W_104440_20200901_1.vrl
 |        `--VR2W-RLD_104440_20200901_1.vrl
 |        `--VR2W_123456_20200901_1.vrl
 |        `--VR2W-RLD_123456_20200901_1.vrl
```
    
 - Do this for each field download day until final downloads for the season are complete (this will typically be by June 30th of each year, depending on conditions).
 
## 2. Copy all vrls from a given field download range into a new subfolder of `YB_Vemco/` using the `ybt::copy_vrls()` function.

Name the new subfolder with the range of download dates covered and a "_vrls" suffix.  This will result in a directory structure like the following:

```{r, echo=TRUE, eval=FALSE}
YB_Vemco
 |-- YBDetections_2020_0901-2021_0523_vrls
    `--VR2W_104440_20200901_1.vrl
    `--VR2W_123456_20200901_1.vrl
    `--VR2W_104440_20210523_1.vrl
    `--VR2W_123456_20210523_1.vrl
 |-- YBFieldDownloads
 |    |-- Field2020_0901
 |        `--Field2020_0901.vdb
 |        `--VR2W_104440_20200901_1.vrl
 |        `--VR2W-RLD_104440_20200901_1.vrl
 |        `--VR2W_123456_20200901_1.vrl
 |        `--VR2W-RLD_123456_20200901_1.vrl
 |    |-- Field2021_0523
 |        `--Field2020_0901.vdb
 |        `--VR2W_104440_20210523_1.vrl
 |        `--VR2W-RLD_104440_20210523_1.vrl
 |        `--VR2W_123456_20210523_1.vrl
 |        `--VR2W-RLD_123456_20210523_1.vrl

```

## 3. Create a new _dc.vdb in that subfolder, following the naming conventions below.

 - Create _dc.vdb (with correct naming convention; see below) and place within the corresponding download range's _dc subfolder (which was created in the previous step).  Example:

```{r, echo=TRUE, eval=FALSE}
YB_Vemco
 |-- YBDetections_2020_0901-2021_0523_vrls
        |-- YBDetections_2020_0901-2021_0523_dc
            `-- YBDetections_2020_0901-2021_0523_dc.vdb
 |-- YBFieldDownloads
```


 - Edit VUE options to "copy .vrl files" into the correct _dc subfolder so that all edited .vrl files will end up in the _dc subfolder during import.

## 4. Drift-correct and export all detections for detection year
   
 - Import raw field .vrls with the VRL editor tool and drift-correct upon import; accept the default appendix to filenames ("_edited.vrl").
 
 - If any errors are reported in the import log, save the import log as a .txt file with and _error tag, i.e.: "YBDetections_2020_0901-2021_0523_dc_Log_todaysdate_error.txt", within the _dc subfolder.  If we run into weird issues during the analysis, this is a good starting point to help us trace problematic or missing detection files.  NO need to save the log if everything imports successfully.

 - Export drift-corrected detections as one .csv into the same _dc subfolder.
      - naming convention: "YBDetections_2020_0901-2021_0523_dc.csv" (specific date range will vary).
  
##   5. Format the detection .csv(s) in R with the `ybt::format_VUE_detections()` function


#### The formatting function does the following:
 - Formats column headings
 - Formats dates/times (POSIXct, PST, yyyy-mm-dd hh:mm:ss) and saves them as text in preparation for database appending
 - Checks for simultaneous detections within receivers and tags; discard duplicates (there really shouldn't be any if VUE is functioning properly)


## 6. Append the resulting detections dataframe to SQLite database

This is done using the `ybt::ybt_db_append()` function; see `vignettes('Database', package = 'ybt')`

